---
title: "Exploratory analysis of penalties"
format:
  html:
    embed-resources: true
editor: visual
---

# Data description

Where the data is from? What are the observations (note there can be several types -- players, plays, penalties), and how many of them? What's the time period covered? What's the geography covered? Is there missing data?

```{r,echo=F,message=F}
library(tidyverse)
library(here)
players <- read.csv(here::here("data/player_info.csv"))
penalties <- read.csv(here::here("data/game_penalties.csv"))
player_plays <- read.csv(here::here("data/plays_players.csv"))
```

## Number of Observations

```{r}
nrow(players)
nrow(penalties)
nrow(player_plays)
```

Since the penalties dataset has more observations than players, this shows that there are players who have committed more than one penalty.

## Example 1: heights

An example of plotting height:

```{r}
ggplot(data = players, aes(x = height_cm)) +
  geom_histogram() +
  theme_minimal() +
  labs(x="Height (cm)", y = "Count")
```

Looks weird. What is there space between bars?

```{r}
players %>% 
  count(height_cm)
```

It looks like height is ordered categorical variable, not a continuous one. Weird!

## Example 2: national origins of players

48% come from Canada, more than from the US. Interesting! I wonder if nationality can predict penalties.

```{r}
players %>% 
  count(nationality) %>% 
  mutate(pct = round(n / sum(n),2)) %>% 
  arrange(-n)
```

## primary position

Are players that play certain positions more prone to causing penalties?

```{r}
players %>%
  count(primaryPosition) 
```

## Dominant Hand

Do right-handed players have more penalties than left-handed players?

```{r}
players %>%
  count(shootsCatches)
```

More lefties than righties is interesting.

## Types of Penalties

```{r}
penalties %>%
  count(penaltySeverity)
```

## Penalty Minutes

```{r}
ggplot(data = penalties, aes(x = penaltyMinutes)) +
  geom_histogram() +
  theme_minimal() +
  labs(x="Penalty Minutes", y = "Occurences")
```

A large majority of the penalties only had two minute punishments, while there are few occurrences of 5 and 10 minute penalties. Could this show deterrence? Are players who get a 10 minute penalty unlikely to commit penalties in the future? Could look at the timeline of this.

```{r}
ggplot(data = player_plays, aes(x = player_id)) +
  geom_histogram(binwidth = 0.5) +
  theme_minimal() +
  labs(x="Player ID", y = "Penalty Count")
```

There is definitely a range- some players have very few penalties whereas other players have hundreds.

```{r}
player_plays %>%
  count(player_id)
```
