---
title: "Exploratory analysis of penalties"
format:
  html:
    embed-resources: true
editor: visual
---

# Data description

Where the data is from? What are the observations (note there can be several types -- players, plays, penalties), and how many of them? What's the time period covered? What's the geography covered? Is there missing data?

```{r,echo=F,message=F}
library(tidyverse)
library(here)
players <- read.csv(here::here("data/player_info.csv"))
```

## Example 1: heights

An example of plotting height:

```{r}
ggplot(data = players, aes(x = height_cm)) +
  geom_histogram() +
  theme_minimal() +
  labs(x="Height (cm)", y = "Count")
```

Looks weird. What is there space between bars?

```{r}
players %>% 
  count(height_cm)
```

It looks like height is ordered categorical variable, not a continuous one. Weird!

## Example 2: national origins of players

48% come from Canada, more than from the US. Interesting! I wonder if nationality can predict penalties.

```{r}
players %>% 
  count(nationality) %>% 
  mutate(pct = round(n / sum(n),2)) %>% 
  arrange(-n)
```
